---
title: 'HIV-SN in children'
subtitle: 'Data analysis: Risk factors for motor dysfunction'
author: Natalie Benjamin, Joanne Potterton, and Peter Kamerman 
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    theme: yeti
    highlight: haddock
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include = FALSE}
# Load libraries
library(tidyverse)
library(stringr)
library(coin)
library(pander)

# knitr options
knitr::opts_chunk$set(echo = TRUE,
               warning = FALSE, 
               message =  FALSE, 
               cache = TRUE,
               fig.path = './figures/sn-risk/',
               tidy = FALSE)
```


```{r data_import}
# Import data
data <- read_csv('./data/cleaned_data.csv')

# Trim df to required variables
data <- data %>%
    select(ID,
           Neuropathy,
           Sex,
           WAZ,
           HAZ,
           CD4,
           Viral.load,
           MABC.zone,
           ends_with('percentile'),
           starts_with('BPNS'),
           starts_with('Other'),
           -ends_with('RL'),
           -ends_with('R'),
           -ends_with('L')) %>%
    rename(MABC.dexterity = MABC.manual.dex.percentile,
           MABC.throwing = MABC.throw.percentile,
           MABC.balance = MABC.balance.percentile,
           MABC.total = MABC.total.percentile,
           Current.pain = BPNS.current.pain,
           Previous.pain = BPNS.prev.pain,
           Numbness = BPNS.numbness,
           Pins.n.needles = BPNS.pin.n.needles,
           Cramping = Other.symp.cramping,
           Itch = Other.symp.itching,
           Painful.cold = Other.symp.painful.cold)
```

## Quick look
```{r quick_look}
# 'Top-n-tail' data
head(data)
tail(data)

# Check structure
glimpse(data)
```

## Predictors

### Neuropathy

**NOTE:** 

- _Risk for symptomatic SN were not explored because of the small sample size of SN+._

```{r motor_sn}
############################################################
#                                                          #
#                 Motor performance in SN+                 #
#                                                          #
############################################################
# Print summary table
data %>%
    select(ID, Neuropathy, starts_with('MABC'), -MABC.zone) %>%
    gather(key = Key, value = value, starts_with('MABC'), -ID, -Neuropathy) %>%
    mutate(Key = str_replace(Key, 'MABC.', ''),
           Key = str_to_title(Key)) %>%
    group_by(Neuropathy, Key) %>%
    summarise(Q25 = round(quantile(value, 0.25)),
              median = round(median(value)),
              Q75 = round(quantile(value, 0.75))) %>%
    arrange(Key, desc(Neuropathy)) %>%
    select(Key, Neuropathy, Q25, median, Q75) %>%
    pander(caption = 'Relationship between neuropathy and MABC variables',
           justify = 'llrrr', emphasize.italics.rows = c(2, 4, 6, 8))
# Summary plot
data %>%
    select(ID, Neuropathy, starts_with('MABC'), -MABC.zone) %>%
    gather(key = Key, value = value, starts_with('MABC'), -ID, -Neuropathy) %>%
    mutate(Key = str_replace(Key, 'MABC.', ''),
           Key = str_to_title(Key)) %>%
    ggplot(data = .) +
    aes(x = Key, y = value, 
        colour = Neuropathy, fill = Neuropathy) +
    geom_boxplot(alpha = 0.6) +
    labs(title = 'Relationship between neuropathy and MABC variables',
         x = 'MABC variable',
         y = 'Percentile')

############################################################
#                                                          #
#                         Balance                          #
#                                                          #
############################################################
# Process data
sn_balance <- data %>%
    select(Neuropathy, MABC.balance)
# Plot
qplot(data = sn_balance, 
      y = MABC.balance, x = Neuropathy, 
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100))
# Stats
wilcox_test(MABC.balance ~ factor(Neuropathy),
            data = sn_balance,
            distribution = 'exact',
            conf.int = TRUE)

############################################################
#                                                          #
#                        Dexterity                         #
#                                                          #
############################################################
# Process data
sn_dexterity <- data %>%
    select(Neuropathy, MABC.dexterity)
# Plot
qplot(data = sn_dexterity, 
      y = MABC.dexterity, x = Neuropathy, 
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100))
# Stats
wilcox_test(MABC.dexterity ~ factor(Neuropathy),
            data = sn_dexterity,
            distribution = 'exact',
            conf.int = TRUE)

############################################################
#                                                          #
#                         Throwing                         #
#                                                          #
############################################################
# Process data
sn_throw <- data %>%
    select(Neuropathy, MABC.throwing)
# Plot
qplot(data = sn_throw, 
      y = MABC.throwing, x = Neuropathy, 
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100))
# Stats
wilcox_test(MABC.throwing ~ factor(Neuropathy),
            data = sn_throw,
            distribution = 'exact',
            conf.int = TRUE)
############################################################
#                                                          #
#                          Total                           #
#                                                          #
############################################################
# Process data
sn_total <- data %>%
    select(Neuropathy, MABC.total)
# Plot
qplot(data = sn_total, 
      y = MABC.total, x = Neuropathy, 
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100))
# Stats
wilcox_test(MABC.total ~ factor(Neuropathy),
            data = sn_total,
            distribution = 'exact',
            conf.int = TRUE)
```

### Other

**NOTE:** 

- _No influence of SN on motor function, therefore collapsed across SN groups._ 

- _Assessed the effect of weight for age, CD4 T-cell count, and viral load_

- _Effect of each 'Other' variable assessed against all MABC variables._

- _If bivariate plot (continious variables) showed no relationship, no correlation performed._ 

```{r motor_other}
############################################################
#                                                          #
#                          Weight                          #
#                                                          #
############################################################

# Balance #
###########
# Process data
waz_balance <- data %>%
    select(WAZ, MABC.balance) %>%
    filter(complete.cases(.))
# Plot
qplot(data = waz_balance, 
      y = MABC.balance, x = WAZ, 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# dexterity #
###########
# Process data
waz_dexterity <- data %>%
    select(WAZ, MABC.dexterity) %>%
    filter(complete.cases(.))
# Plot
qplot(data = waz_dexterity, 
      y = MABC.dexterity, x = WAZ, 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# Throwing #
###########
# Process data
waz_throwing <- data %>%
    select(WAZ, MABC.throwing) %>%
    filter(complete.cases(.))
# Plot
qplot(data = waz_throwing, 
      y = MABC.throwing, x = WAZ, 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# total #
###########
# Process data
waz_total <- data %>%
    select(WAZ, MABC.total) %>%
    filter(complete.cases(.))
# Plot
qplot(data = waz_total, 
      y = MABC.total, x = WAZ, 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

############################################################
#                                                          #
#                          CD4                             #
#                                                          #
############################################################

# Balance #
###########
# Process data
cd4_balance <- data %>%
    select(CD4, MABC.balance) %>%
    filter(complete.cases(.))
# Plot
qplot(data = cd4_balance, 
      y = MABC.balance, x = CD4, 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# dexterity #
###########
# Process data
cd4_dexterity <- data %>%
    select(CD4, MABC.dexterity) %>%
    filter(complete.cases(.))
# Plot
qplot(data = cd4_dexterity, 
      y = MABC.dexterity, x = CD4, 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# Throwing #
###########
# Process data
cd4_throwing <- data %>%
    select(CD4, MABC.throwing) %>%
    filter(complete.cases(.))
# Plot
qplot(data = cd4_throwing, 
      y = MABC.throwing, x = CD4, 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# total #
###########
# Process data
cd4_total <- data %>%
    select(CD4, MABC.total) %>%
    filter(complete.cases(.))
# Plot
qplot(data = cd4_total, 
      y = MABC.total, x = CD4, 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

############################################################
#                                                          #
#                  (log10)Viral load                       #
#                                                          #
############################################################

# Balance #
###########
# Process data
viral.load_balance <- data %>%
    select(Viral.load, MABC.balance) %>%
    filter(complete.cases(.))
# Plot
qplot(data = viral.load_balance, 
      y = MABC.balance, x = log10(Viral.load), 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# dexterity #
###########
# Process data
viral.load_dexterity <- data %>%
    select(Viral.load, MABC.dexterity) %>%
    filter(complete.cases(.))
# Plot
qplot(data = viral.load_dexterity, 
      y = MABC.dexterity, x = log10(Viral.load), 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# Throwing #
###########
# Process data
viral.load_throwing <- data %>%
    select(Viral.load, MABC.throwing) %>%
    filter(complete.cases(.))
# Plot
qplot(data = viral.load_throwing, 
      y = MABC.throwing, x = log10(Viral.load), 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship

# total #
###########
# Process data
viral.load_total <- data %>%
    select(Viral.load, MABC.total) %>%
    filter(complete.cases(.))
# Plot
qplot(data = viral.load_total, 
      y = MABC.total, x = log10(Viral.load), 
      geom = c('point', 'smooth'), 
      ylim = c(0, 100))
# No relationship
```

## Session infomation
```{r session_info, echo = F, warning = F, message = F}
pander(sessionInfo())
```

