---
title: 'HIV-SN in children'
subtitle: 'Data analysis: Risk factors for motor dysfunction'
author: Peter Kamerman 
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    theme: yeti
    highlight: haddock
    code_folding: show
    toc: true
    toc_depth: 5
    toc_float: true
    df_print: paged
---

```{r setup, include = FALSE}
# Load libraries
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(stringr)
library(coin)
library(pander)

# knitr options
knitr::opts_chunk$set(echo = TRUE,
               warning = FALSE, 
               message =  FALSE)

# Import data
data <- read_csv('./data/cleaned_data.csv')

# Trim df to required variables
data <- data %>%
    select(ID,
           Neuropathy,
           Neuropathy_pin,
           Sex,
           WAZ,
           HAZ,
           CD4,
           Viral.load,
           MABC.zone,
           ends_with('percentile'),
           starts_with('BPNS'),
           starts_with('Other'),
           matches('\\.RL$'),
           -matches('\\.L$'),
           -matches('\\.R$')) %>%
    rename(dexterity = MABC.manual.dex.percentile,
           throwing = MABC.throw.percentile,
           balance = MABC.balance.percentile,
           total = MABC.total.percentile,
           reduced.vibration = BPNS.reduced.vibration.RL,
           reduced.reflex = BPNS.absent.reflex.RL,
           reduced.pinprick = Reduced.pin.prick.RL,
           current.pain = BPNS.current.pain,
           previous.pain = BPNS.prev.pain,
           numbness = BPNS.numbness,
           pins.n.needles = BPNS.pin.n.needles,
           cramping = Other.symp.cramping,
           itch = Other.symp.itching,
           painful.cold = Other.symp.painful.cold)
```

## Quick look at the data
```{r quick_look}
# 'Top-n-tail' data
head(data)
tail(data)

# Check structure
glimpse(data)
```

****

## Predictors

### Neuropathy
**NOTE:** 

- _Risk associated with symptomatic SN were not explored because of the small sample size._

#### HIV-SN (pin-prick)
**Case definition of SN:**  

- _Brief Peripheral Neuropathy Screen (BPNS) + pin-prick_

- _At least one bilateral sign (reduced/absent vibration sense, absent ankle-jerk reflexes, reduced/absent pin-prick)_  

##### Summary plot
```{r pin_summary}
# Summary plot
data %>%
    select(ID, Neuropathy_pin, dexterity, throwing, balance, total) %>%
    gather(key = Key, value = value, dexterity, throwing, balance, total,
           -ID, -Neuropathy_pin) %>%
    rename(`Neuropathy (pin-prick)` = Neuropathy_pin) %>%
    ggplot(data = .) +
    aes(x = Key, y = value, 
        colour = `Neuropathy (pin-prick)`, fill = `Neuropathy (pin-prick)`) +
    geom_boxplot(alpha = 0.6) +
    scale_fill_brewer(type = 'qual', palette = 'Dark2') +
    scale_colour_brewer(type = 'qual', palette = 'Dark2') +
    labs(title = 'Relationship between neuropathy and MABC variables',
         x = 'MABC variable',
         y = 'Performance (Percentile)') +
    theme_minimal(base_size = 14)
```

##### Balance
```{r pin_balance}
# Process data
sn_balance <- data %>%
    select(Neuropathy_pin, balance)

# Plot
qplot(data = sn_balance, 
      y = balance, x = Neuropathy_pin, 
      xlab ='Neuropathy (pin-prick)', ylab = 'Performance (Percentile)',
      main = 'Effect of SN (pin-prick) on MABC balance',
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(balance ~ factor(Neuropathy_pin),
            data = sn_balance,
            distribution = 'exact',
            conf.int = TRUE)
```

##### Dexterity
```{r pin_dexterity}
# Process data
sn_dexterity <- data %>%
    select(Neuropathy_pin, dexterity)

# Plot
qplot(data = sn_dexterity, 
      y = dexterity, x = Neuropathy_pin, 
      xlab ='Neuropathy (pin-prick)', ylab = 'Performance (Percentile)',
      main = 'Effect of SN (pin-prick) on MABC manual dexterity',
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(dexterity ~ factor(Neuropathy_pin),
            data = sn_dexterity,
            distribution = 'exact',
            conf.int = TRUE)
```

##### Throwing
```{r pin_throwing}
# Process data
sn_throw <- data %>%
    select(Neuropathy_pin, throwing)

# Plot
qplot(data = sn_throw, 
      y = throwing, x = Neuropathy_pin, 
      xlab ='Neuropathy (pin-prick)', ylab = 'Performance (Percentile)',
      main = 'Effect of SN (pin-prick) on MABC throwing',
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(throwing ~ factor(Neuropathy_pin),
            data = sn_throw,
            distribution = 'exact',
            conf.int = TRUE)
```

##### total
```{r pin_total}
# Process data
sn_total <- data %>%
    select(Neuropathy_pin, total)

# Plot
qplot(data = sn_total, 
      y = total, x = Neuropathy_pin, 
      xlab ='Neuropathy (pin-prick)', ylab = 'Performance (Percentile)',
      main = 'Effect of SN (pin-prick) on MABC overall',
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(total ~ factor(Neuropathy_pin),
            data = sn_total,
            distribution = 'exact',
            conf.int = TRUE)
```

****

#### HIV-SN (BPNS)
**Case definition of SN:**  

- _Brief Peripheral Neuropathy Screen (BPNS)_

- _At least one bilateral sign (reduced/absent vibration sense, absent ankle-jerk reflexes)_  

##### Summary plot
```{r summary}
# Summary plot
data %>%
    select(ID, Neuropathy, dexterity, throwing, balance, total) %>%
    gather(key = Key, value = value, dexterity, throwing, balance, total,
           -ID, -Neuropathy) %>%
    rename(`Neuropathy (BPNS)` = Neuropathy) %>%
    ggplot(data = .) +
    aes(x = Key, y = value, 
        colour = `Neuropathy (BPNS)`, fill = `Neuropathy (BPNS)`) +
    geom_boxplot(alpha = 0.6) +
    scale_fill_brewer(type = 'qual', palette = 'Dark2') +
    scale_colour_brewer(type = 'qual', palette = 'Dark2') +
    labs(title = 'Relationship between neuropathy and MABC variables',
         x = 'MABC variable',
         y = 'Performance (Percentile)') +
    theme_minimal(base_size = 14)
```

##### Balance
```{r balance}
# Process data
sn_balance <- data %>%
    select(Neuropathy, balance)

# Plot
qplot(data = sn_balance, 
      y = balance, x = Neuropathy, 
      xlab ='Neuropathy (BPNS)', ylab = 'Performance (Percentile)',
      main = 'Effect of SN (BPNS) on MABC balance',
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(balance ~ factor(Neuropathy),
            data = sn_balance,
            distribution = 'exact',
            conf.int = TRUE)
```

##### Dexterity
```{r dexterity}
# Process data
sn_dexterity <- data %>%
    select(Neuropathy, dexterity)

# Plot
qplot(data = sn_dexterity, 
      y = dexterity, x = Neuropathy, 
      xlab ='Neuropathy (BPNS)', ylab = 'Performance (Percentile)',
      main = 'Effect of SN (BPNS) on MABC manual dexterity',
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(dexterity ~ factor(Neuropathy),
            data = sn_dexterity,
            distribution = 'exact',
            conf.int = TRUE)
```

##### Throwing
```{r throwing}
# Process data
sn_throw <- data %>%
    select(Neuropathy, throwing)

# Plot
qplot(data = sn_throw, 
      y = throwing, x = Neuropathy, 
      xlab ='Neuropathy (BPNS)', ylab = 'Performance (Percentile)',
      main = 'Effect of SN (BPNS) on MABC throwing',
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(throwing ~ factor(Neuropathy),
            data = sn_throw,
            distribution = 'exact',
            conf.int = TRUE)
```

##### total
```{r total}
# Process data
sn_total <- data %>%
    select(Neuropathy, total)

# Plot
qplot(data = sn_total, 
      y = total, x = Neuropathy, 
      xlab ='Neuropathy (BPNS)', ylab = 'Performance (Percentile)',
      main = 'Effect of SN (BPNS) on MABC overall',
      geom = c('boxplot', 'jitter'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(total ~ factor(Neuropathy),
            data = sn_total,
            distribution = 'exact',
            conf.int = TRUE)
```

****

### Other risk factors

**NOTE:** 

- _No influence of SN on motor function, therefore collapsed across SN groups._ 

- _Assessed the effect of weight for age, CD4 T-cell count, and viral load_

- _Effect of each 'Other' variable assessed against all MABC variables._

- _If bivariate plot (continious variables) showed no relationship, no correlation performed._ 
##### Weight
```{r weight}
#################
#    Balance    #
#################
# Process data
waz_balance <- data %>%
    select(WAZ, balance) %>%
    filter(complete.cases(.))

# Plot
qplot(data = waz_balance, 
      y = balance, x = WAZ,
      xlab ='Weight for age (z-score)', ylab = 'Performance (Percentile)',
      main = 'Relationship between weight and MABC balance',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

###################
#    Dexterity    #
###################
# Process data
waz_dexterity <- data %>%
    select(WAZ, dexterity) %>%
    filter(complete.cases(.))

# Plot
qplot(data = waz_dexterity, 
      y = dexterity, x = WAZ, 
      xlab ='Weight for age (z-score)', ylab = 'Performance (Percentile)',
      main = 'Relationship between weight and MABC manual dexterity',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

##################
#    Throwing    #
##################
# Process data
waz_throwing <- data %>%
    select(WAZ, throwing) %>%
    filter(complete.cases(.))

# Plot
qplot(data = waz_throwing, 
      y = throwing, x = WAZ, 
      xlab ='Weight for age (z-score)', ylab = 'Performance (Percentile)',
      main = 'Relationship between weight and MABC throwing',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

###############
#    Total    #
###############
# Process data
waz_total <- data %>%
    select(WAZ, total) %>%
    filter(complete.cases(.))

# Plot
qplot(data = waz_total, 
      y = total, x = WAZ, 
      xlab ='Weight for age (z-score)', ylab = 'Performance (Percentile)',
      main = 'Relationship between weight and MABC total',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED
```

##### CD4 T-cell count
```{r cd4}
#################
#    Balance    #
#################
# Process data
CD4_balance <- data %>%
    select(CD4, balance) %>%
    filter(complete.cases(.))

# Plot
qplot(data = CD4_balance, 
      y = balance, x = CD4,
      xlab ='CD4 T-cell count', ylab = 'Performance (Percentile)',
      main = 'Relationship between CD4 and MABC balance',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

###################
#    Dexterity    #
###################
# Process data
CD4_dexterity <- data %>%
    select(CD4, dexterity) %>%
    filter(complete.cases(.))

# Plot
qplot(data = CD4_dexterity, 
      y = dexterity, x = CD4, 
      xlab ='CD4 T-cell count', ylab = 'Performance (Percentile)',
      main = 'Relationship between CD4 and MABC manual dexterity',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

##################
#    Throwing    #
##################
# Process data
CD4_throwing <- data %>%
    select(CD4, throwing) %>%
    filter(complete.cases(.))

# Plot
qplot(data = CD4_throwing, 
      y = throwing, x = CD4, 
      xlab ='CD4 T-cell count', ylab = 'Performance (Percentile)',
      main = 'Relationship between CD4 and MABC throwing',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

###############
#    Total    #
###############
# Process data
CD4_total <- data %>%
    select(CD4, total) %>%
    filter(complete.cases(.))

# Plot
qplot(data = CD4_total, 
      y = total, x = CD4, 
      xlab ='CD4 T-cell count', ylab = 'Performance (Percentile)',
      main = 'Relationship between CD4 and MABC total',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED
```

##### (log10)Viral load
```{r viral_load}
#################
#    Balance    #
#################
# Process data
VL_balance <- data %>%
    select(Viral.load, balance) %>%
    filter(complete.cases(.))

# Plot
qplot(data = VL_balance, 
      y = balance, x = Viral.load,
      xlab ='log10(viral load)', ylab = 'Performance (Percentile)',
      main = 'Relationship between log10(viral load) MABC balance',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

###################
#    Dexterity    #
###################
# Process data
VL_dexterity <- data %>%
    select(Viral.load, dexterity) %>%
    filter(complete.cases(.))

# Plot
qplot(data = VL_dexterity, 
      y = dexterity, x = Viral.load, 
      xlab ='log10(viral load)', ylab = 'Performance (Percentile)',
      main = 'Relationship between log10(viral load) MABC manual dexterity',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

##################
#    Throwing    #
##################
# Process data
VL_throwing <- data %>%
    select(Viral.load, throwing) %>%
    filter(complete.cases(.))

# Plot
qplot(data = VL_throwing, 
      y = throwing, x = Viral.load, 
      xlab ='log10(viral load)', ylab = 'Performance (Percentile)',
      main = 'Relationship between log10(viral load) MABC throwing',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED

###############
#    Total    #
###############
# Process data
VL_total <- data %>%
    select(Viral.load, total) %>%
    filter(complete.cases(.))

# Plot
qplot(data = VL_total, 
      y = total, x = Viral.load, 
      xlab ='log10(viral load)', ylab = 'Performance (Percentile)',
      main = 'Relationship between log10(viral load) MABC total',
      geom = c('point', 'smooth'), 
      ylim = c(0, 100)) +
    theme_minimal(base_size = 14)

# NO RELATIONSHIP, NO FURTHER ANALYSIS PERFORMED
```

## Session infomation
```{r session_info, echo = F, warning = F, message = F}
pander(sessionInfo())
```

