---
title: 'HIV-SN in children'
subtitle: 'Descriptive analyses: Neuropathy'
author: Peter Kamerman 
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    theme: yeti
    highlight: haddock
    code_folding: show
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
---

```{r setup, include = FALSE}
# Load librarieslibrary(dplyr)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(stringr)
library(pander)
library(boot)

# knitr options
knitr::opts_chunk$set(echo = TRUE,
               warning = FALSE, 
               message =  FALSE, 
               fig.align = 'center')

# Import data
data <- read_csv('./data/cleaned_data.csv')

# Trim df to prevalence data
data <- select(data,
               ID,
               starts_with('BPNS'), 
               starts_with('Other'),
               ends_with('.RL'),
               -ends_with('.R'),
               -ends_with('.L'),
               Symptoms, 
               Neuropathy,
               Neuropathy_pin,
               Symp.Neuropathy,
               Symp.Neuropathy_pin)
```

## Quick look at the data
```{r quick_look}
# 'Top-n-tail' data
head(data)
tail(data)

# Check structure
glimpse(data)
```

****

## Prevalence of HIV-SN (pin-prick)
**Case definition of SN:**  

- _Brief Peripheral Neuropathy Screen (BPNS) + pin-prick_

- _At least one bilateral sign (reduced/absent vibration sense, absent ankle-jerk reflexes, reduced/absent pin-prick)_  

**Case definition of symptomatic SN:** 

- _As above, plus at least one bilateral symptom (pain/aching/burning, tingling/pins-and-needles, numbness)_

### Boot functions
```{r pin_boot_functions}
############################################################
#                                                          #
#                      Boot functions                      #
#                                                          #
############################################################
# Use package 'boot' to calculate bootstrap 95%CI

## Any SN
sn_prev <- function(data, i) {
    dat <- data[i, ]
    len <- dat %>%
        select(Neuropathy_pin) %>%
        filter(Neuropathy_pin == 'Yes') %>%
        nrow()
    len2 <- nrow(dat)
    prev <- 100 * len / len2
    prev
}

## Symptomatic SN
symp.sn_prev <- function(data, i) {
    dat <- data[i,]
    len <- dat %>%
        select(Symp.Neuropathy_pin) %>%
        filter(Symp.Neuropathy_pin == 'Yes') %>%
        nrow()
    len2 <- nrow(dat)
    prev <- 100 * len / len2
    return(prev)
}
```

### Prevalence summaries
#### SN prevalence (whole cohort) 
```{r pin_sn_prevalence}
# Counts
sn_point <- data %>%  
    group_by(Neuropathy_pin) %>%  
    summarise(N = n(), Percent = 100 * (n() / nrow(data)))

pander(sn_point[ , 1:2], 
       caption = 'SN vs SN-free (count)', justify = 'lr')

# Bootstrap 95% CI
sn_boot <- boot(data = data, 
              statistic = sn_prev, 
              R = 1000)

sn_bootCI <- boot.ci(sn_boot, type = 'basic')

# Summary table
sn_tab <- data_frame(Group = 'Neuropathy_pin',
                     `Point prevalence` = round(sn_point$Percent[2], 1),
                     `Lower 95% CI limit` = round(sn_bootCI$basic[4], 1),
                     `Upper 95% CI limit` = round(sn_bootCI$basic[5], 1)) 
pander(sn_tab[ , 2:4], 
       caption = 'Prevalence of SN (%)', justify = 'rrr')
```

#### Symptomatic SN+ (SN+ cohort only)
```{r pin_symp_sn_prevalence}
# Subset out the SN+ participants
data.sn_only <- data %>%
    filter(Neuropathy_pin == 'Yes')

# Counts
symp.sn_point <- data.sn_only %>%  
    group_by(Symp.Neuropathy_pin) %>%  
    summarise(N = n(), Percent = 100 * (n() / nrow(data.sn_only)))

pander(symp.sn_point[ , 1:2], 
       caption = 'Symptomatic SN (SN+ only, count)', justify = 'lr')

# Bootstrap 95% CI
symp.sn_boot <- boot(data = data.sn_only, 
                     statistic = symp.sn_prev, 
                     R = 1000)

symp.sn_bootCI <- boot.ci(symp.sn_boot, type = 'basic')

# Summary table
symp.sn_tab <- data_frame(Group = 'Symptomatic HIV-SN',
                          `Point prevalence` = round(symp.sn_point$Percent[2], 1),
                          `Lower 95% CI limit` = round(symp.sn_bootCI$basic[4], 1),
                          `Upper 95% CI limit` = round(symp.sn_bootCI$basic[5], 1))
pander(symp.sn_tab[ , 2:4], 
       caption = 'Prevalence of Symptomatic SN (SN+ only, %)', justify = 'rrr')
```

### Characteristics of symptomatic SN
```{r pin_symptoms}
############################################################
#                                                          #
#     Not very informative because of the small number     #
#         (n = 12) of children with symptomatic SN          #
#                                                          #
############################################################
# Filter out only those with symptomatic SN (primary definition)
symptomatic_sn <- data %>%
    filter(Symp.Neuropathy_pin == 'Yes')

# Extract symptoms
symptoms_sn <- symptomatic_sn %>%
    select(BPNS.numbness, BPNS.prev.pain, BPNS.current.pain, BPNS.pin.n.needles,
           Other.symp.itching, Other.symp.cramping, Other.symp.painful.cold) %>%
    rename(Numbness = BPNS.numbness, 
           `Previous pain` = BPNS.prev.pain,
           `Current pain` = BPNS.current.pain, 
           `Pins and needles` = BPNS.pin.n.needles,
           Itch = Other.symp.itching, 
           Cramping = Other.symp.cramping, 
           `Painful cold` = Other.symp.painful.cold)

# Table point estimates of symptom prevalences (%)
symptom_tab <- symptoms_sn %>%
    summarise_each(funs(round(sum(. == 'Yes', na.rm = TRUE)))) %>%
    gather(key = key,
           value = n) %>%
    mutate(percent = round(n / nrow(symptomatic_sn) * 100)) %>%
    arrange(desc(n))

# Table
pander(symptom_tab, 
       caption = 'Prevalence (%) of symptoms in 12 children with symptomatic HIV-SN', 
       justify = 'lrr', split.tables = Inf)
```

### Characteristics of SN signs
```{r pin_signs}
# Filter out only those with SN
sn_signs <- data %>%
    filter(Neuropathy_pin == 'Yes')

# Extract signs
signs_sn <- sn_signs %>%
    select(BPNS.absent.reflex.RL, BPNS.reduced.vibration.RL, Reduced.pin.prick.RL) %>%
    rename(`Absent reflexes` = BPNS.absent.reflex.RL, 
           `Absent/Reduced vibration sense` = BPNS.reduced.vibration.RL,
           `Absent/reduced pin-prick sense` = Reduced.pin.prick.RL)

# Table point estimates of signs prevalences (%)
signs_tab <- signs_sn %>%
    summarise_each(funs(round(sum(. == 'Yes', na.rm = TRUE)))) %>%
    gather(key = key,
           value = n) %>%
    mutate(percent = round(n / nrow(signs_sn) * 100)) %>%
    arrange(desc(n))

# Table
pander(signs_tab, 
       caption = 'Prevalence (%) of signs in 34 children with HIV-SN', 
       justify = 'lrr', split.tables = Inf)
```

****

## Prevalence of HIV-SN (BPNS)
**Case definition of SN:**  

- _Brief Peripheral Neuropathy Screen (BPNS) only_

- _At least one bilateral sign (reduced/absent vibration sense, absent ankle-jerk reflexes; NB: no pin-prick)_  

**Case definition of symptomatic SN:** 

- _As above, plus at least one bilateral symptom (pain/aching/burning, tingling/pins-and-needles, numbness)_

### Boot functions
```{r boot_functions}
############################################################
#                                                          #
#                      Boot functions                      #
#                                                          #
############################################################
# Use package 'boot' to calculate bootstrap 95%CI

## Any SN
sn_prev <- function(data, i) {
    dat <- data[i, ]
    len <- dat %>%
        select(Neuropathy) %>%
        filter(Neuropathy == 'Yes') %>%
        nrow()
    len2 <- nrow(dat)
    prev <- 100 * len / len2
    prev
}

## Symptomatic SN
symp.sn_prev <- function(data, i) {
    dat <- data[i,]
    len <- dat %>%
        select(Symp.Neuropathy) %>%
        filter(Symp.Neuropathy == 'Yes') %>%
        nrow()
    len2 <- nrow(dat)
    prev <- 100 * len / len2
    return(prev)
}
```

### Prevalence summaries
#### SN prevalence (whole cohort) 
```{r sn_prevalence}
# Counts
sn_point <- data %>%  
    group_by(Neuropathy) %>%  
    summarise(N = n(), Percent = 100 * (n() / nrow(data)))

pander(sn_point[ , 1:2], 
       caption = 'SN vs SN-free (count)', justify = 'lr')

# Bootstrap 95% CI
sn_boot <- boot(data = data, 
              statistic = sn_prev, 
              R = 1000)

sn_bootCI <- boot.ci(sn_boot, type = 'basic')

# Summary table
sn_tab <- data_frame(Group = 'Neuropathy',
                     `Point prevalence` = round(sn_point$Percent[2], 1),
                     `Lower 95% CI limit` = round(sn_bootCI$basic[4], 1),
                     `Upper 95% CI limit` = round(sn_bootCI$basic[5], 1)) 

pander(sn_tab[ , 2:4], 
       caption = 'Prevalence of SN (%)', justify = 'rrr')
```

#### Symptomatic SN+ (SN+ cohort only)
```{r symp_sn_prevalence}
# Subset out the SN+ participants
data.sn_only <- data %>%
    filter(Neuropathy == 'Yes')

# Counts
symp.sn_point <- data.sn_only %>%  
    group_by(Symp.Neuropathy) %>%  
    summarise(N = n(), Percent = 100 * (n() / nrow(data.sn_only)))

pander(symp.sn_point[ , 1:2], 
       caption = 'Symptomatic SN (SN+ only, count)', justify = 'lr')

# Bootstrap 95% CI
symp.sn_boot <- boot(data = data.sn_only, 
                     statistic = symp.sn_prev, 
                     R = 1000)

symp.sn_bootCI <- boot.ci(symp.sn_boot, type = 'basic')

# Summary table
symp.sn_tab <- data_frame(Group = 'Symptomatic HIV-SN',
                          `Point prevalence` = round(symp.sn_point$Percent[2], 1),
                          `Lower 95% CI limit` = round(symp.sn_bootCI$basic[4], 1),
                          `Upper 95% CI limit` = round(symp.sn_bootCI$basic[5], 1))

pander(symp.sn_tab[ , 2:4], 
       caption = 'Prevalence of Symptomatic SN (SN+ only, %)', justify = 'rrr')
```

### Characteristics of symptomatic SN
```{r symptoms}
############################################################
#                                                          #
#     Not very informative because of the small number     #
#         (n = 5) of children with symptomatic SN          #
#                                                          #
############################################################
# Filter out only those with symptomatic SN (primary definition)
symptomatic_sn <- data %>%
    filter(Symp.Neuropathy == 'Yes')

# Extract symptoms
symptoms_sn <- symptomatic_sn %>%
    select(BPNS.numbness, BPNS.prev.pain, BPNS.current.pain, BPNS.pin.n.needles,
           Other.symp.itching, Other.symp.cramping, Other.symp.painful.cold) %>%
    rename(Numbness = BPNS.numbness, 
           `Previous pain` = BPNS.prev.pain,
           `Current pain` = BPNS.current.pain, 
           `Pins and needles` =BPNS.pin.n.needles,
           Itch = Other.symp.itching, 
           Cramping = Other.symp.cramping, 
           `Painful cold` = Other.symp.painful.cold)

# Table point estimates of symptom prevalences (%)
symptom_tab <- symptoms_sn %>%
    summarise_each(funs(round(sum(. == 'Yes', na.rm = TRUE)))) %>%
    gather(key = key,
           value = n) %>%
    mutate(percent = round(n / nrow(symptomatic_sn) * 100)) %>%
    arrange(desc(n))#/ length(.x) * 100, 1)) 

# Table
pander(symptom_tab, 
       caption = 'Prevalence (%) of symptoms in 5 children with symptomatic HIV-SN', 
       justify = 'lrr', split.tables = Inf)
```

### Characteristics of SN signs
```{r signs}
# Filter out only those with SN
sn_signs <- data %>%
    filter(Neuropathy == 'Yes')

# Extract signs
signs_sn <- sn_signs %>%
    select(BPNS.absent.reflex.RL, BPNS.reduced.vibration.RL) %>%
    rename(`Absent reflexes` = BPNS.absent.reflex.RL, 
           `Absent/Reduced vibration sense` = BPNS.reduced.vibration.RL)

# Table point estimates of signs prevalences (%)
signs_tab <- signs_sn %>%
    summarise_each(funs(round(sum(. == 'Yes', na.rm = TRUE)))) %>%
    gather(key = key,
           value = n) %>%
    mutate(percent = round(n / nrow(signs_sn) * 100)) %>%
    arrange(desc(n))

# Table
pander(signs_tab, 
       caption = 'Prevalence (%) of signs in 18 children with HIV-SN', 
       justify = 'lrr', split.tables = Inf)
```

****

## Session infomation
```{r session_info, echo = F, warning = F, message = F}
pander(sessionInfo())
```
