---
title: 'HIV-SN in children'
subtitle: 'Data analysis: Risk factors for SN'
author: Peter Kamerman 
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    theme: yeti
    highlight: haddock
    code_folding: show
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
---

```{r setup, include = FALSE}
# Load libraries
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(coin)
library(pander)

# knitr options
knitr::opts_chunk$set(echo = TRUE,
               warning = FALSE, 
               message =  FALSE)

# Import data
data <- read_csv('./data/cleaned_data.csv')

# Trim df to required variables
data <- data %>%
    select(ID,
           Neuropathy,
           Neuropathy_pin,
           Age.years,
           Sex,
           WAZ,
           HAZ,
           CD4,
           Viral.load,
           d4T.ever,
           ART.years)
```

## Quick look
```{r quick_look}
# 'Top-n-tail' data
head(data)
tail(data)

# Check structure
glimpse(data)
```

****

## Prevalence of HIV-SN (pin-prick)
**Case definition of SN:**  

- _Brief Peripheral Neuropathy Screen (BPNS) + pin-prick_

- _At least one bilateral sign (reduced/absent vibration sense, absent ankle-jerk reflexes, reduced/absent pin-prick)_  

**Case definition of symptomatic SN:** 

- _As above, plus at least one bilateral symptom (pain/aching/burning, tingling/pins-and-needles, numbness)_

### Predictors
**NOTE:** 

- _Predictors for symptomatic SN not explored because of the very small sample size._ 

**Examined:**  

- _Age, sex, height for age, weight for age, CD4 T-cell count, and viral load_

#### Age
```{r pin_age}
# Select data and filter out incomplete cases
age <- data %>%
    select(Age.years, Neuropathy_pin) %>%
    filter(complete.cases(.))

# Plot
qplot(data = age, x = Neuropathy_pin, y = Age.years, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14) 

# Stats
wilcox_test(Age.years ~ factor(Neuropathy_pin), 
            data = age,
            distribution = 'exact',
            conf.int = TRUE)
```

#### Sex
```{r pin_sex}
# Select data and filter out incomplete cases
sex <- data %>%
    select(Sex, Neuropathy_pin) %>%
    filter(complete.cases(.)) %>%
    table()

# Plot
mosaicplot(sex,
           cex.axis = 1)

# Stats
chisq_test(sex,
           distribution = 'exact')
```

#### Height
```{r pin_height}
# Select data and filter out incomplete cases
haz <- data %>%
    select(HAZ, Neuropathy_pin) %>%
    filter(complete.cases(.))

# Plot
qplot(data = haz, x = Neuropathy_pin, y = HAZ, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

## Stats
wilcox_test(HAZ ~ factor(Neuropathy_pin),
            data = haz,
            distribution = 'exact',
            conf.int = TRUE)
```

#### Age
```{r pin_weight}
# Select data and filter out incomplete cases
waz <- data %>%
    select(WAZ, Neuropathy_pin) %>%
    filter(complete.cases(.))

# Plot
qplot(data = waz, x = Neuropathy_pin, y = WAZ, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(WAZ ~ factor(Neuropathy_pin),
            data = waz,
            distribution = 'exact',
            conf.int = TRUE)
```

#### CD4
```{r pin_cd4}
# Select data and filter out incomplete cases
cd4 <- data %>%
    select(CD4, Neuropathy_pin) %>%
    filter(complete.cases(.))

# Plot
qplot(data = cd4, x = Neuropathy_pin, y = CD4, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

## Stats
wilcox_test(CD4 ~ factor(Neuropathy_pin),
            data = cd4,
            distribution = 'exact',
            conf.int = TRUE)
```

#### Viral load
```{r pin_viral_load}
# Select data and filter out incomplete cases
vl <- data %>%
    select(Viral.load, Neuropathy_pin) %>%
    filter(complete.cases(.)) %>%
    mutate(log10_viral.load = log10(Viral.load))

# Plot
qplot(data = vl, x = Neuropathy_pin, y = log10_viral.load, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(log10_viral.load ~ factor(Neuropathy_pin), 
            data = vl, 
            distribution = 'exact',
            conf.int = TRUE)
```

#### D4T use ever
```{r pin_d4t_ever}
# Select data and filter out incomplete cases
d4t <- data %>%
    select(d4T.ever, Neuropathy_pin) %>%
    filter(complete.cases(.)) %>%
    table()
# Plot
mosaicplot(d4t,
           cex.axis = 1)
# Stats
chisq_test(d4t,
           distribution = 'exact',
           conf.int = TRUE)
```

#### ART duration
```{r pin_art_duration}
# Select data and filter out incomplete cases
art.duration <- data %>%
    select(ART.years, Neuropathy_pin) %>%
    filter(complete.cases(.))

# Plot
qplot(data = art.duration, x = Neuropathy_pin, y = ART.years,
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(ART.years ~ factor(Neuropathy_pin),
            data = art.duration,
            distribution = 'exact',
            conf.int = TRUE)
```

****

## Prevalence of HIV-SN (BPNS)
**Case definition of SN:**  

- _Brief Peripheral Neuropathy Screen (BPNS) only_

- _At least one bilateral sign (reduced/absent vibration sense, absent ankle-jerk reflexes)_  

**Case definition of symptomatic SN:** 

- _As above, plus at least one bilateral symptom (pain/aching/burning, tingling/pins-and-needles, numbness)_

### Predictors
**NOTE:** 

- _Predictors for symptomatic SN not explored because of the very small sample size._  

**Examined:**  

- _Age, sex, height for age, weight for age, CD4 T-cell count, and viral load_

#### Age
```{r age}
# Select data and filter out incomplete cases
age <- data %>%
    select(Age.years, Neuropathy) %>%
    filter(complete.cases(.))

# Plot
qplot(data = age, x = Neuropathy, y = Age.years, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(Age.years ~ factor(Neuropathy), 
            data = age,
            distribution = 'exact',
            conf.int = TRUE)
```

#### Sex
```{r sex}
# Select data and filter out incomplete cases
sex <- data %>%
    select(Sex, Neuropathy) %>%
    filter(complete.cases(.)) %>%
    table()

# Plot
mosaicplot(sex,
           cex.axis = 1)

# Stats
chisq_test(sex,
           distribution = 'exact')
```

#### Height
```{r height}
# Select data and filter out incomplete cases
haz <- data %>%
    select(HAZ, Neuropathy) %>%
    filter(complete.cases(.))

# Plot
qplot(data = haz, x = Neuropathy, y = HAZ, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

## Stats
wilcox_test(HAZ ~ factor(Neuropathy),
            data = haz,
            distribution = 'exact',
            conf.int = TRUE)
```

#### Age
```{r weight}
# Select data and filter out incomplete cases
waz <- data %>%
    select(WAZ, Neuropathy) %>%
    filter(complete.cases(.))

# Plot
qplot(data = waz, x = Neuropathy, y = WAZ, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(WAZ ~ factor(Neuropathy),
            data = waz,
            distribution = 'exact',
            conf.int = TRUE)
```

#### CD4
```{r cd4}
# Select data and filter out incomplete cases
cd4 <- data %>%
    select(CD4, Neuropathy) %>%
    filter(complete.cases(.))

# Plot
qplot(data = cd4, x = Neuropathy, y = CD4, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

## Stats
wilcox_test(CD4 ~ factor(Neuropathy),
            data = cd4,
            distribution = 'exact',
            conf.int = TRUE)
```

#### Viral load
```{r viral_load}
# Select data and filter out incomplete cases
vl <- data %>%
    select(Viral.load, Neuropathy) %>%
    filter(complete.cases(.)) %>%
    mutate(log10_viral.load = log10(Viral.load))

# Plot
qplot(data = vl, x = Neuropathy, y = log10_viral.load, 
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(log10_viral.load ~ factor(Neuropathy), 
            data = vl, 
            distribution = 'exact',
            conf.int = TRUE)
```

#### D4T use ever
```{r d4t_ever}
# Select data and filter out incomplete cases
d4t <- data %>%
    select(d4T.ever, Neuropathy) %>%
    filter(complete.cases(.)) %>%
    table()

# Plot
mosaicplot(d4t,
           cex.axis = 1)

# Stats
chisq_test(d4t,
           distribution = 'exact',
           conf.int = TRUE) 
```

#### ART duration
```{r art_duration}
# Select data and filter out incomplete cases
art.duration <- data %>%
    select(ART.years, Neuropathy) %>%
    filter(complete.cases(.))

# Plot
qplot(data = art.duration, x = Neuropathy, y = ART.years,
      geom = c('boxplot', 'jitter')) +
    theme_minimal(base_size = 14)

# Stats
wilcox_test(ART.years ~ factor(Neuropathy),
            data = art.duration,
            distribution = 'exact',
            conf.int = TRUE)
```

****

## Session infomation
```{r session_info, echo = F, warning = F, message = F}
pander(sessionInfo())
```

