---
title: 'HIV-SN in children'
author: Natalie Benjamin, Joanne Potterton, and Peter Kamerman 
date: "`r format(Sys.time(), '%d %B %Y')`"
output: github_document
---

```{r setup, include = FALSE}
# Load libraries
library(readr)
library(pander)

# knitr options
knitr::opts_chunk$set(echo = FALSE,
               warning = FALSE, 
               message =  FALSE, 
               cache = FALSE,
               fig.path = './figures/',
               tidy = FALSE)
```

## Project overview

## Ethics clearance
The study was approved by the Human Ethics Research Committee of the University of the Witwatersrand, South Africa (clearance certificate: M120767).

## Questions  
1. Prevalence of:
    - HIV-associated sensory neuropathy (SN)  
    - Symptomatic SN  
2. Characteristics of symptomatic SN 
3. Predictors of SN/symptomatic SN (secondary analysis)
4. Motor development
    - Descriptive data on motor development  
    - Does SN influence motor development  
    - Do other disease-related factors influence motor development (secondary analysis)  

## Code book
```{r code_book, echo = FALSE}
# Import code book 
codebook <- read_csv('./data/code_book.csv')

# Table
pander(codebook, 
       style = 'rmarkdown', 
       justify = 'left',
       split.tables = 'Inf')
```